Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping

### 2025-06-04 CaseStudies/Example3_4_SimulationandEstimationofLoopSettlingTimes/article/article.qmd
Finishing examples.

---
title: Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping
subtitle: Case Study - Example 3.4 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design and analysis of a fractional-N frequency synthesizer based on a charge pump phase-locked loop (CP-PLL) architecture. The system comprises a voltage-controlled oscillator (VCO), a charge pump, a reference oscillator, and a second-order loop filter. Key parameters such as the loop bandwidth, natural frequency, and damping factor are used to guide the filter design. The settling time is estimated for different frequency steps, and the system’s response is evaluated for both linear and nonlinear operating regions. The methodology follows a continuous-time approach.
keywords: 
    - Fractional-N PLL
    - Linearity Limit
    - Cycle Slipping
    - Nonlinear Dynamics
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

from IPython.display import display, Markdown

def format(value, decimals=2):
    abs_value = abs(value)
    fmt = f".{decimals}f"

    if abs_value >= 1e9:
        result = f"{value / 1e9:{fmt}}\\,\\text{{G}}"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:{fmt}}\\,\\text{{M}}"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:{fmt}}\\,\\text{{k}}"
    elif abs_value >= 1:
        result = f"{value * 1e0:{fmt}}\\,"
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:{fmt}}\\,\\text{{m}}"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:{fmt}}\\,\\mu"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:{fmt}}\\,\\text{{n}}"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:{fmt}}\\,\\text{{p}}"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:{fmt}}\\,\\text{{f}}"
    else:
        result = f"{value:.0e}\\,"

    return Markdown(f"${result}$")

@@

```{python}
#| echo: false
import math
from IPython.display import display, Markdown

def format_with_pi_latex(value):
    pi_multiple = value / math.pi
    abs_pi_multiple = abs(pi_multiple)

    if abs_pi_multiple >= 1e9:
        result = f"{pi_multiple / 1e9:.2f}\\pi\\,\\text{{G}}"
    elif abs_pi_multiple >= 1e6:
        result = f"{pi_multiple / 1e6:.2f}\\pi\\,\\text{{M}}"
    elif abs_pi_multiple >= 1e3:
        result = f"{pi_multiple / 1e3:.2f}\\pi\\,\\text{{k}}"
    elif abs_pi_multiple >= 1:
        result = f"{pi_multiple:.2f}\\pi"
    elif abs_pi_multiple >= 1e-3:
        result = f"{pi_multiple * 1e3:.2f}\\pi\\,\\text{{m}}"
    elif abs_pi_multiple >= 1e-6:
        result = f"{pi_multiple * 1e6:.2f}\\pi\\,\\mu"
    elif abs_pi_multiple >= 1e-9:
        result = f"{pi_multiple * 1e9:.2f}\\pi\\,\\text{{n}}"
    elif abs_pi_multiple >= 1e-12:
        result = f"{pi_multiple * 1e12:.2f}\\pi\\,\\text{{p}}"
    elif abs_pi_multiple >= 1e-15:
        result = f"{pi_multiple * 1e15:.2f}\\pi\\,\\text{{f}}"
    else:
        result = f"{pi_multiple:.0e}\\pi"

    return Markdown(f"${result}$")

# Exemplo de uso:

import math

f0=3.7e9
f = 4.3e9
tunning_range = f-f0
step=1e6
fref = 40e6
I_CP=2*math.pi*100e-6
V_sup=3
f_3dB=150e3
omega_3dB=f_3dB*2*math.pi

Kvco=tunning_range/V_sup
Kvco_ang=Kvco*2*math.pi

f_30M=30e6
f_300M=300e6

Ktheta=I_CP/(2*math.pi)

fvco=4e9
N = fvco/fref
```

# Introduction

A `{python} format(f0)`Hz–`{python} format(f)`Hz synthesizer with a step size of `{python} format(step)`Hz is required. A `{python} format(fref)`Hz crystal oscillator, a charge pump with `{python} format(I_CP)`A output current, and a VCO (operating from a `{python} format(V_sup)`V supply) are available. The objective is to design a fractional-N synthesizer with a loop bandwidth of `{python} format(f_3dB)`Hz using these components. The settling time of the loop must be estimated for frequency steps of `{python} format(f_30M)`Hz and `{python} format(f_300M)`Hz. Simulation results should then be analyzed and compared. This example is adapted from @rogers2006IntegratedCircuitDesign.

# Design Methodology

## Voltage-Controlled Oscillator (VCO)

The VCO operates with a supply voltage of `{python} format(V_sup)`V and offers a tuning range of `{python} format(tunning_range)`Hz. Therefore, the VCO gain constant is given by $K_\text{VCO}=$ `{python} format(Kvco)`Hz/V. The nominal frequency is arbitrarily selected as $f_\text{VCO}=$ `{python} format(fvco)`Hz. With this nominal frequency, the corresponding division ratio is $N=$ `{python} N`.

## Charge pump

Given the known charge pump current, the charge pump gain constant can be expressed as $K_\theta=$ `{python} format(Ktheta)`A/rad.

## Loop filter

import math

fref_f3dB=fref/f_3dB
```

Since the ratio of the reference frequency to the loop bandwidth is $f_\text{ref}/f_{3\text{dB}}=$ `{python} round(fref_f3dB, 0)`, i.e., $f_\text{ref}\gg f_{3\text{dB}}$, a continuous-time analysis is appropriate for this system.

The loop filter design begins with the calculation of the natural frequency:

$$
\omega_n=\frac{\omega_{3\text{dB}}}{\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}}
$$

@@

zeta = 0.707

omega_n=omega_3dB/math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
f_n=omega_n/(2*math.pi)
```

Using the chosen damping factor $\zeta=$ `{python} zeta`, the natural frequency is $f_n=$ `{python} format(f_n)`Hz.

The value of the first capacitor, $C_1$, is determined by:

$$
C_1=\frac{IK_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

@@

C_1=I_CP*Kvco_ang/(2*math.pi*N*omega_n**2)

This results in $C_1=$ `{python} format(C_1)`F.

The first resistor in the second-order loop filter is given by:

$$
R_1=2\zeta\sqrt{\frac{2\pi\cdot N}{IK_\text{VCO}C_1}}
$$

import math

R_1=2*zeta*math.sqrt((2*math.pi*N)/(I_CP*Kvco_ang*C_1))
C_2=C_1/10
```

Thus, the value of the resistor is $R_1=$ `{python} format(R_1)`$\Omega$. The second capacitor is typically chosen to be one-tenth of $C_1$, yielding $C_2=$ `{python} format(C_2)`F.

## System Testing and Performance Estimation

@@

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 2 * math.pi
```

@@

The maximum tolerable input frequency step for this configuration is $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s, equivalent to `{python} format(step_m_lin)`Hz.

@@

f_i_30M = f_30M/N
f_i_30M_ang = f_i_30M*2*math.pi
f_i_300M = f_300M/N
f_i_300M_ang = f_i_300M*2*math.pi
```

omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
t_s=omega_n__t_s/omega_n
```

Given output frequency steps of `{python} format(f_30M)`Hz and `{python} format(f_300M)`Hz, the corresponding reference frequency steps are `{python} format(f_i_30M)`Hz and `{python} format(f_i_300M)`Hz, respectively.

The first frequency step lies within the linear range and thus must settle within a predictable time interval, computed via the $\omega_nt_s$ relationship. The settling time is found to be $t_s=$ `{python} format(t_s)`s.

In contrast, the `{python} format(f_i_300M)`Hz frequency step exceeds the linear region, resulting in cycle slipping. The acquisition time in this nonlinear case is estimated using the equation:

$$
T_s=\frac{\Delta\omega}{\pi\omega_n^2}
$$

Ts = f_i_300M_ang/(math.pi*omega_n**2)
Total_ts=Ts+t_s
```

Accordingly, the acquisition time is $T_s=$ `{python} format(Ts)`s. An additional `{python} format(t_s)`s is required for phase locking, leading to a total locking time of `{python} format(Total_ts)`s for a `{python} format(f_i_300M)`Hz step.

<!--```{python}
#| echo: false
import sympy as sp
from IPython.display import display, Markdown

# Variáveis simbólicas
s = sp.symbols('s')
I_sym, Kvco_sym, R1_sym, C1_sym, N_sym = sp.symbols('I Kvco R1 C1 N')

# Constante K = (I * Kvco) / (2π)
K_sym = (I_sym * Kvco_sym) / (2 * sp.pi)

# Numerador e denominador da função de transferência
num = K_sym / C1_sym * (1 + R1_sym * C1_sym * s)
den = s**2 + (K_sym * R1_sym / N_sym) * s + K_sym / (N_sym * C1_sym)

# Função de transferência
H_s = num / den

# Substitui os valores
H_subs = H_s.subs({I_sym: I_CP, Kvco_sym: Kvco_ang, C1_sym: C_1, R1_sym: R_1, N_sym: N}).evalf(4)

# Converte para LaTeX
latex_str = sp.latex(H_subs)

display(Markdown(f"$$\\frac{{\\varphi_\\text{{no}}(s)}}{{\\varphi_\\text{{nI}}(s)}} = {latex_str}$$"))
```-->

# Conclusion

A fractional-N PLL synthesizer was designed using the provided specifications and components. The loop filter parameters were calculated based on continuous-time modeling, and the loop performance was evaluated under both linear and nonlinear frequency step scenarios. Settling times were estimated analytically. Further validation through simulation is necessary to confirm theoretical predictions.

# References {-}

>>>>

---
title: Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping
subtitle: Case Study - Example 3.4 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    Example 3.4.
keywords: 
    - Fractional-N PLL
    - Linearity Limit
    - Cycle Slipping
    - Nonlinear Dynamics
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

from format_utils import format, format_graph, format_no_markdown, format_with_pi_latex

@@

@@

zeta = 0.707
f_bw = 150e3
omega_bw = 2 * math.pi * f_bw
```

@@

omega_n=omega_bw/math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
```

@@

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = (1.6255) * math.pi
```

@@

@@

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```

import numpy as np

# Parâmetros iniciais
f_r = 40e6 # Hz
omega_r = 2 * np.pi * f_r # rad/s

fvco = f_o = 4e9 # Hz
omega_o = 2 * np.pi * f_o # rad/s
N = 100
Kvco = KVCO = 200e6 # Hz/V
Kvco_ang = 2 * np.pi * Kvco # rad/s/V
I_CP = I = 100e-6
K_theta = Kphase = I / (2 * np.pi) # A/rad
A0 = 6.258
K = A0 * Kphase * KVCO / N
```

from scipy import signal

omega_n = omega_bw / np.sqrt(1 + 2 * zeta**2 + np.sqrt(4 * zeta**4 + 4 * zeta**2 + 2))
f_n = omega_n/2/np.pi
```

# Parâmetros do PLL
C_1 = C1 = K / omega_n**2
C_2 = C2 = C1 / 10
R_1 = R1 = 2 * zeta * omega_n / K
```

import math
import numpy as np

# Análise de resposta em frequência
omega_n_meas = np.sqrt(K / C1)
f_n_meas = omega_n_meas / (2 * np.pi)

zeta_meas = (R1 / 2) * np.sqrt(K * C1)

t_m_meas = 1 / (omega_n_meas * math.sqrt(1 - zeta_meas**2)) * math.atan(math.sqrt(1 - zeta_meas**2) / zeta_meas)

theta_e_m_n_meas = math.sin(omega_n_meas * math.sqrt(1 - zeta_meas**2) * t_m_meas) / \
                   (math.sqrt(1 - zeta_meas**2)) * math.e**(-zeta_meas * omega_n_meas * t_m_meas)

theta_e_m_meas = theta_e_m

omega_bw_meas = omega_n_meas * np.sqrt(1 + 2 * zeta_meas**2 + np.sqrt(4 * zeta_meas**4 + 4 * zeta_meas**2 + 2))
f_bw_meas = omega_bw_meas / (2 * np.pi)

step_m_ang_meas = theta_e_m_meas * omega_n_meas / theta_e_m_n_meas
step_m_lin_meas = step_m_ang_meas / (2 * math.pi)

t_s_meas = math.log(0.01 * math.sqrt(1 - zeta_meas**2)) / (zeta_meas * omega_n_meas) * (-1)

from scipy import signal
from IPython.display import display, Markdown

def format_eng(x):
    if x == 0:
        return "0"
    exponent = int("{:.0e}".format(x).split('e')[1])
    base = x / (10**exponent)
    return f"{base:.2f}e{exponent}"

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)
```

# Numerador e denominador do compensador F(s)
# F(s) = (s*C1*R1 + 1) / (s^2*C1*C2*R1 + s*(C1 + C2))
F_num = [C1 * R1, 1]
F_den = [C1 * C2 * R1, C1 + C2, 0]

F_sys = signal.TransferFunction(F_num, F_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = F_sys.num[0][0] if isinstance(F_sys.num[0], (list, tuple)) else F_sys.num
den = F_sys.den[0][0] if isinstance(F_sys.den[0], (list, tuple)) else F_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-filter-numeric}}"))

```{python}
#| echo: false
# F(s) * K / s = G(s)
G_num = np.polymul(F_num, [K])
G_den = np.polymul(F_den, [1, 0])  # Multiplica por s

G_sys = signal.TransferFunction(G_num, G_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = G_sys.num[0][0] if isinstance(G_sys.num[0], (list, tuple)) else G_sys.num
den = G_sys.den[0][0] if isinstance(G_sys.den[0], (list, tuple)) else G_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$G(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-open-loop-numeric}}"))
```

```{python}
#| echo: false
# H(s) = G(s) / (1 + G(s))
H_num = G_num
H_den = np.polyadd(G_den, G_num)

# Criando o sistema de tempo contínuo
H_sys = signal.TransferFunction(H_num, H_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = H_sys.num[0][0] if isinstance(H_sys.num[0], (list, tuple)) else H_sys.num
den = H_sys.den[0][0] if isinstance(H_sys.den[0], (list, tuple)) else H_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-closed-loop-numeric}}"))
```

# Agora temos o denominador completo
# O_sys = H_sys * N
# Para isso, O_sys = (H_num * N) / (H_den)
num_O = np.convolve(H_sys.num, [N])
den_O = np.convolve(H_sys.den, [1])

# Criar a função de transferência final
O_sys = signal.TransferFunction(num_O, den_O)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = O_sys.num[0][0] if isinstance(O_sys.num[0], (list, tuple)) else O_sys.num
den = O_sys.den[0][0] if isinstance(O_sys.den[0], (list, tuple)) else O_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$O(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-output-phase-numeric}}"))
```

# Kvco/s => [Kvco] / [1, 0]
Kvco_over_s = ([Kvco], [1, 0])

# F_sys => ([1], [1, 1]) já está
# Multiplicando numeradores e denominadores via convolução
num_den = np.convolve([1/N], Kvco_over_s[0])
num_den = np.convolve(num_den, F_sys.num)
num_den = np.convolve(num_den, [K_theta])
num_den = np.convolve(num_den, [A0])

den_den = np.convolve(Kvco_over_s[1], F_sys.den)

# Agora temos o denominador completo
# E_sys = H_sys / (den)
# Para isso, E_sys = (H_num * den_den) / (H_den * den_num)
num_E = np.convolve(H_sys.num, den_den)
den_E = np.convolve(H_sys.den, num_den)

# Criar a função de transferência final
E_sys = signal.TransferFunction(num_E, den_E)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = E_sys.num[0][0] if isinstance(E_sys.num[0], (list, tuple)) else E_sys.num
den = E_sys.den[0][0] if isinstance(E_sys.den[0], (list, tuple)) else E_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$E(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-phase-error-numeric}}"))
```

# Vc_sys = H_sys / ((1/N)*(Kvco/s))
#Kvco_over_s = ([Kvco], [1, 0])
Kvco_over_s = ([Kvco], [1])
den_num = np.convolve([1/N], Kvco_over_s[0])
den_den = Kvco_over_s[1]

num_Vc = np.convolve(H_sys.num, den_den)
den_Vc = np.convolve(H_sys.den, den_num)

Vc_sys = signal.TransferFunction(num_Vc, den_Vc)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = Vc_sys.num[0][0] if isinstance(Vc_sys.num[0], (list, tuple)) else Vc_sys.num
den = Vc_sys.den[0][0] if isinstance(Vc_sys.den[0], (list, tuple)) else Vc_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$V_C(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-vctrl-vco-numeric}}"))
```

import numpy as np

omega_r_safe = omega_n*10
f_r_safe = omega_r_safe/2/np.pi

omega_r_max = omega_n * 2 * np.pi * zeta
f_r_max = omega_r_max/2/np.pi
```

sim_stop_time = 50e-6
sim_points = 10000
step_amplitude = f_r
```

```{python}
#| echo: false
#| label: fig-step-freq-h
#| fig-cap: "Frequency step response of the closed-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Valor final da resposta
final_value = y_out[-1]
tol = 0.01 * final_value  # 1% da resposta final

# Encontrar o tempo em que a resposta entra e permanece na faixa de ±1%
ts_graphical = 0
within_bounds = np.abs(y_out - u) <= tol
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback frequency')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference frequency')
plt.axhline(final_value + tol, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(final_value - tol, color='red', linestyle='--', linewidth=0.8)
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling time: {ts_graphical*1e6:.5f} µs')
plt.xlabel('Time (µs)')
plt.ylabel('Frequency (Hz)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-ramp-phase-h
#| fig-cap: "Phase response to a ramp input in the closed-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference Phase')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-ramp-phase-h-o
#| fig-cap: "Comparison of phase responses to a ramp input in two different functions."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)
t_out, y_out2, _ = signal.lsim(O_sys, U=u, T=t)

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, y_out2, label='Output Phase')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference Phase')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-ramp-phase-step-freq-h
#| fig-cap: "Phase response to a ramp with an abrupt frequency change (frequency step)."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Frequências
f_base = omega_r  # 10 MHz
f_step = step_m_ang_meas  # step de 10 MHz
t_step = 20e-6  # instante do step

# Entrada em fase: integral da frequência
phi_in = np.where(
    t < t_step,
    f_base * t,
    f_base * t + f_step * (t - t_step)
)

# Simulação da resposta
t_out, y_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

### --- Gráfico ---
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, phi_in, linestyle=':', color='black', label='Reference Phase')

plt.axvline(t_step * 1e6, color='purple', linestyle='--', linewidth=1.2, label='Step Frequency Instant')

plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-ramp-phase-h-phase-error
#| fig-cap: "Response to a ramp input with phase error."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Entrada: rampa em fase (step de 10 MHz em frequência)
# Fase = ∫ frequência ⇒ rampa
freq_step = omega_r  # Step de 10 MHz em frequência
phi_in = freq_step * t  # Rampa de fase

# Simulação da resposta do sistema à rampa
t_out, phi_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

# Erro de seguimento
tracking_error = phi_in - phi_out

# Plotagem
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, phi_in, linestyle=':', color='black', label='Reference Phase')
plt.plot(t_out * 1e6, phi_out, label='Feedback Phase')
plt.plot(t_out * 1e6, tracking_error, label='Phase Error')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-step-freq-e
#| fig-cap: "Phase error response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0.01 * max_value  # 1% do valor de pico (centrado em zero)

# Novo cálculo do tempo de acomodação (settling time)
within_bounds = np.abs(y_out) <= tol
ts_graphical_peak = 0
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical_peak = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Phase Error')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='red', linestyle='--', linewidth=0.8, label=f'±1% Tolerance')
plt.axhline(-tol, color='red', linestyle='--', linewidth=0.8)

# Marcação do novo tempo de acomodação
plt.axvline(ts_graphical_peak * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling Time: {ts_graphical_peak*1e6:.3f} μs')

plt.xlabel('Time (µs)')
plt.ylabel('Phase Error (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-step-freq-e-norm
#| fig-cap: "Normalized phase error response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 8/omega_n_meas, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

y_out *= omega_n_meas

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * omega_n_meas, y_out, label='Step Response')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value: {max_value:.3f}')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='black', linestyle=':')

plt.xlabel('Time (µs)')
plt.ylabel('Phase Error norm. (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-step-freq-vc
#| fig-cap: "Control voltage response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import TransferFunction, step

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Resposta ao degrau
t, vc_response = step(Vc_sys, T=t)
vc_response *= step_m_lin_meas

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N - f_o) / KVCO
#final_value = 0.5
tol = 0.01 * final_value  # 1% de tolerância
upper_bound = final_value + tol
lower_bound = final_value - tol

# Encontrar settling time: busca o primeiro instante em que o sinal entra e permanece dentro da faixa de tolerância
within_bounds = np.logical_and(vc_response >= lower_bound, vc_response <= upper_bound)

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 3*150  # ~5% da simulação

ts_graphical = 0
for i in range(len(within_bounds) - window_size):
    if np.all(within_bounds[i:i + window_size]):
        ts_graphical = t[i]
        break

# Plotagem
plt.figure(figsize=(8, 6))
plt.plot(t * 1e6, vc_response, label='Voltage Control')
plt.axhline(upper_bound, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(lower_bound, color='red', linestyle='--', linewidth=0.8)

if ts_graphical is not None:
    plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2,
                label=f'Settling time: {ts_graphical*1e6:.2f} µs')

plt.xlabel('Time (µs)')
plt.ylabel('Voltage Control (V)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-step-freq-vc-norm
#| fig-cap: "Normalized control voltage response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import TransferFunction, step

# Tempo de simulação
t = np.linspace(0, 8/omega_n_meas, sim_points)

# Resposta ao degrau
t, vc_response = step(Vc_sys, T=t)
vc_response *= Kvco / N

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N - f_o) / KVCO
final_value *= KVCO / (step_m_lin_meas * N)
#final_value = 0.5

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 50  # ~5% da simulação

# Plotagem
plt.figure(figsize=(8, 6))
plt.plot(t * omega_n_meas, vc_response, label='Voltage Control')
plt.axhline(final_value, color='black', linestyle=':', label='Reference Voltage')

plt.xlabel('Tempo (µs)')
plt.ylabel('Voltage Control norm. (V)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-bode-g
#| fig-cap: "Bode plot of the open-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(G_sys)

# Plotando o diagrama de Bode
plt.figure(figsize=(8, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-bode-h
#| fig-cap: "Bode plot of the closed-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(H_sys)
